---
title: "Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an accompanying repository for the paper *"Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls''* by Pavla Krotka, Martin Posch, Mohamed Gewily, Günter Höglinger, and Marta Bofill Roig. It contains the code to reproduce all simulations and figures, as well as the case study presented in the paper.
<!-- "[Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls](https://arxiv.org)". -->

The repository is structured as follows:

- Folder **simulations**:

    - *NCC_FreqModels_simscript.R*: This script contains the code to reproduce all simulations included in the paper. The results for the individual scenarios are then saved in the subfolder *results*.
    - *NCC_FreqModels_figures.Rmd*: This file contains the code to create all figures presented in the paper (Section 4) and the supplementary material (Section C). The figures are saved in the subfolder *figures* in a .pdf, .png and .tiff formats.

- Folder **case_studies**:

    - Subfolder **PSP**:

        - *NCC_FreqModels_case_study_PSP.Rmd*: This file includes the code used for illustrating the use of the methods from the paper on the data from the ABBV-8E12 trial, as shown in Section 5. It also contains the code to reproduce Figure 13, which is saved in the subfolder *figures*. For data privacy reasons the dataset used for the illustration could not be uploaded.
        - *NCC_FreqModels_case_study_PSP_synthetic.Rmd*: This file includes the code used for illustrating the use of the methods from the paper on synthetic data that mimic the data from the ABBV-8E12 trial.
        

    - Subfolder **FLAIR**:
        - *NCC_FreqModels_case_study_PSP.Rmd*: This file includes the code used for illustrating the use of the methods from the paper on simulated data based on the FLAIR trial, as shown in the supplementary material (Section D). It also contains the code to reproduce Figure S12, which is saved in the subfolder *figures*.
    
## NCC R-package

The version of the `NCC` [R-package](https://pavlakrotka.github.io/NCC/) that was used for the simulation study is included in this repository.

This version is labeled as Release 1.3 on [GitHub](https://github.com/pavlakrotka/NCC) and can be installed by running the following code:

```{r, message=FALSE, warning=FALSE, results=FALSE, eval=FALSE}
# install.packages("devtools") 
devtools::install_github("pavlakrotka/NCC@v1.3", force = TRUE, build_vignettes = TRUE)
```

------------------------------------------------------------------------

**Funding**

This research was funded in whole or in part by the Austrian Science Fund (FWF) [ESP 442 ESPRIT-Programm].

P.K., M.P., M.G., and G.H. received funding from the European Joint Programme on Rare Diseases (Improve-PSP).