---
title: "Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls: Case study"
author: "Pavla Krotka, Martin Posch, Marta Bofill Roig"
date: "2024"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show 
    theme: flatly
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(collapse = TRUE,
               echo = TRUE,
               comment = "#>",
               warning = FALSE,
               message = FALSE,
               fig.align = "center",
               out.width = "100%")
```

```{r, include=FALSE}
library(ggpubr)
library(scales)
library(kableExtra)
library(latex2exp)
library(tidyverse)
library(splines)
library(lmerTest)
library(spaMM)
library(viridis)

#devtools::install_github("pavlakrotka/NCC", build = TRUE, force=T)
library(NCC)
set.seed(717)
```


# Introduction

This file contains all code to reproduce the case study presented in Section 5 of the paper *"Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls"* by Pavla Krotka, Martin Posch and Marta Bofill Roig.


# Data preparation

```{r}
improve_data <- read_csv("final_baseline.csv")
```


```{r}
data_mod <- improve_data %>%
  filter(Study=="Abbvie", Item_number==35, Visit %in% c(1, 11, 14, 18)) %>%
  select(ID, response = DV, TRT, baseline_date, TIME, Treatment_number, Visit, Item_number) %>%
  mutate(treatment = as.factor(case_when(Treatment_number==1 ~ 1,
                                         Treatment_number==2 ~ 2,
                                         Treatment_number==3 ~ 0))) %>%
  arrange(baseline_date) %>%
  mutate(recruitment = as.numeric(baseline_date-min(baseline_date)+1),
         period = ifelse(baseline_date<as.Date("2018-06-01"), 1, 2),
         treatment_mod = as.factor(ifelse((period==1 & treatment==2), NA, as.numeric(treatment)-1)))

data_mod$response <- as.numeric(data_mod$response)
data_mod$TRT <- factor(data_mod$TRT, levels = c("Abbvie-Placebo", "Abbvie-ABBV-8E12 2000 MG", "Abbvie-ABBV-8E12 4000 MG"))
data_mod$period <- as.factor(data_mod$period)

data_mod_wide <- data_mod %>%
  select(-TIME) %>%
  pivot_wider(names_from = Visit, names_prefix = "response_visit_", values_from = response) %>%
  group_by(ID) %>%
  mutate(baseline_change_11 = response_visit_11-response_visit_1,
         baseline_change_14 = response_visit_14-response_visit_1,
         baseline_change_18 = response_visit_18-response_visit_1,
         cal_time = factor(ceiling(recruitment/90), levels = c(1:9)))

data_case_study <- data_mod_wide %>%
  select(ID, recruitment, baseline_date, response_visit_1, treatment = treatment_mod, period, cal_time) %>%
  na.omit()
```



# Abbvie trial data - baseline measurement

## Sample sizes per treatment arm

```{r}
table(data_case_study$treatment)
```

## Sample size per period

```{r}
table(data_case_study$period)
```


## Sample size per treatment arm and period

```{r}
table(data_case_study$treatment, data_case_study$period)
```

## Sample sizes per calendar time interval

```{r}
table(data_case_study$cal_time)
```


## Sample sizes per treatment arm and calendar time interval

```{r}
table(data_case_study$treatment, data_case_study$cal_time)
```


## Time trend

```{r}
test_trend <- lm(response_visit_1 ~ treatment + recruitment, data_case_study)
summary(test_trend)
```


```{r}
ggplot(data_case_study) +
  geom_point(aes(recruitment, response_visit_1, color = treatment)) +
  geom_line(aes(recruitment, fitted(test_trend), color = treatment), size = 1) +
  labs(x = "Trial duration in days", y = "PSPRS score at baseline", color = "Treatment:") +
  scale_color_viridis(discrete = T) +
  geom_vline(xintercept = 90*c(0:9), linetype = "dashed", color = "gray40") +
  geom_vline(xintercept = 495, linetype = "dashed", color = "darkred", size = 1) +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("figures/casestudy_trialdata.png", width = 7, height = 5)
ggsave("figures/casestudy_trialdata.pdf", width = 7, height = 5)
```

# Analysis approaches

## Regression model - period adjustment

```{r}
# Regression model - period adjustment
mod_lm_per <- lm(response_visit_1 ~ treatment + period, data_case_study)
summary(mod_lm_per)
```

## Regression model - calendar time adjustment

```{r}
# Regression model - calendar time adjustment
mod_lm_cal <- lm(response_visit_1 ~ treatment + cal_time, data_case_study)
summary(mod_lm_cal)
```
## Mixed model - period adjustment

```{r}
# Mixed model - period adjustment
mod_mix_per <- lmer(response_visit_1 ~ treatment + (1 | period), data_case_study)
summary(mod_mix_per)
```

## Mixed model - calendar time adjustment

```{r}
# Mixed model - calendar time adjustment
mod_mix_cal <- lmer(response_visit_1 ~ treatment + (1 | cal_time), data_case_study)
summary(mod_mix_cal)
```

## Mixed model (AR1) - period adjustment

```{r}
# Mixed model - period adjustment
mod_mix_ar1_per <- fitme(response_visit_1 ~ treatment + AR1(1 | period), data_case_study)

summary.HLfit(mod_mix_ar1_per)
res <- summary.HLfit(mod_mix_ar1_per, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_per, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["treatment2", "t-value"]), eff_df)) # p-value
```

## Mixed model (AR1) - calendar time adjustment

```{r}
# Mixed model - calendar time adjustment
mod_mix_ar1_cal <- fitme(response_visit_1 ~ treatment + AR1(1 | cal_time), data_case_study)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

#pt(res$beta_table["treatment2", "t-value"], eff_df)*2
#2*pt(-abs(res$beta_table["treatment2", "t-value"]), eff_df)
2*(1-pt(abs(res$beta_table["treatment2", "t-value"]), eff_df)) # p-value
```

## Spline regression - period adjustment

```{r}
# Spline regression - period adjustment
mod_splines_per <- lm(response_visit_1 ~ treatment + bs(recruitment, knots = c(495), degree = 3), data_case_study)
summary(mod_splines_per)
```

## Spline regression - calendar time adjustment

```{r}
# Spline regression - calendar time adjustment
cal_time_start <- c()
for (i in unique(data_case_study$cal_time)) {
  cal_time_start <- c(cal_time_start, max(data_case_study[data_case_study$cal_time==i,]$recruitment))
  }

mod_splines_cal <- lm(response_visit_1 ~ treatment + bs(recruitment, knots = cal_time_start, degree = 3), data_case_study)
summary(mod_splines_cal)
```

## Pooled analysis

```{r}
# Pooled analysis
mod_pool <- lm(response_visit_1 ~ treatment, data_case_study %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)
```

## Separate analysis

```{r}
# Separate analysis
mod_sep <- lm(response_visit_1 ~ treatment, data_case_study %>% filter(treatment %in% c(0, 2), period==2))
summary(mod_sep)
```
