---
title: "Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls"
subtitle: "Case study: FLAIR trial"
author: "Pavla Krotka, Martin Posch, Mohamed Gewily, Günter Höglinger, Marta Bofill Roig"
date: "2024"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    code_folding: show 
    theme: flatly
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(collapse = TRUE,
               echo = TRUE,
               comment = "#>",
               warning = FALSE,
               message = FALSE,
               fig.align = "center",
               out.width = "100%")
```

```{r, warning=FALSE, message=FALSE, include=FALSE}
library(tidyverse)
library(kableExtra)
library(splines)
library(lmerTest)
library(spaMM)

# devtools::install_github("pavlakrotka/NCC@v1.3", force = TRUE)
library(NCC)
```


# Introduction

This file contains all code to reproduce the case study based on the FLAIR trial presented in Section 5.2 of the paper *"Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls"* by Pavla Krotka, Martin Posch, Mohamed Gewily, Günter Höglinger, Marta Bofill Roig. The results obtained from the analyses below are presented in Table 3, Table S1, and Figure S11.


# Data simulation

```{r}
# Function to simulate data from the FLAIR trial

sim_flair <- function(lambda, trend_pattern, theta = c(-log(0.44), -log(0.13))){
  
  # treatments: FCR-0, IR-1, IV-2
  
  period1 <- sample(c(rep(0, 316), rep(1, 317)))
  period2 <- sample(c(rep(0, 69), rep(1, 69), rep(2, 69)))
  period3 <- sample(c(rep(0, 194), rep(2, 191)))
  
  trt <- c(period1, period2, period3)
  
  per <- c(rep(1, length(period1)), rep(2, length(period2)), rep(3, length(period3)))
  
  per_trend <- c(rep(1, length(period1)), rep(2, length(period2)+length(period3)))
  
  j0 <- which(trt==0)
  j1 <- which(trt==1)
  j2 <- which(trt==2)
  
  if (trend_pattern=="linear") {
    trend <- linear_trend(1:length(trt), lambda = lambda, sample_size = c(1, length(trt)))
  }
  
  if (trend_pattern=="stepwise") {
    trend <- sw_trend(cj = per_trend, lambda = lambda)
  }
  
  
  means <- c()
  
  means[j0] <- 0
  means[j1] <- theta[1]
  means[j2] <- theta[2]
  
  response <- rnorm(length(trt), means, 1)
  
  # Add time trend
  
  response[j0] <- response[j0] + trend[j0]
  response[j1] <- response[j1] + trend[j1]
  response[j2] <- response[j2] + trend[j2]
  
  unit_size <- 100
  
  flair_data <- data.frame(j = 1:length(trt),
                           treatment = trt,
                           period = per,
                           response = response,
                           cal_time = rep(c(1:ceiling((length(trt)/unit_size))), each = unit_size)[1:length(trt)])
  
  flair_data
}
```

```{r}
# Simulation of 4 settings with different time trends patterns and strengths (using same simulation runs to distinguish changes due to the added time trend)

set.seed(132)
flair_data_lin_0.25 <- sim_flair(lambda = 0.25, trend_pattern = "linear")

set.seed(132)
flair_data_lin_1.5 <- sim_flair(lambda = 1.5, trend_pattern = "linear")

set.seed(132)
flair_data_lin_5 <- sim_flair(lambda = 5, trend_pattern = "linear")

set.seed(132)
flair_data_step_0.25 <- sim_flair(lambda = 0.25, trend_pattern = "stepwise")

set.seed(132)
flair_data_step_1.5 <- sim_flair(lambda = 1.5, trend_pattern = "stepwise")

set.seed(132)
flair_data_step_5 <- sim_flair(lambda = 5, trend_pattern = "stepwise")
```


## Sample sizes per treatment and period

```{r}
table(flair_data_lin_0.25$treatment, flair_data_lin_0.25$period)
```

```{r}
rowSums(table(flair_data_lin_0.25$treatment, flair_data_lin_0.25$period))
```


## Total sample size

```{r}
nrow(flair_data_lin_0.25)
```



# Considered approaches

## Regression model - period adjustment

### Linear trend, $\lambda=0.25$

```{r}
mod_lm_per <- lm(response ~ as.factor(treatment) + as.factor(period), flair_data_lin_0.25)
summary(mod_lm_per)
```

### Linear trend, $\lambda=1.5$

```{r}
mod_lm_per <- lm(response ~ as.factor(treatment) + as.factor(period), flair_data_lin_1.5)
summary(mod_lm_per)
```

### Linear trend, $\lambda=5$

```{r}
mod_lm_per <- lm(response ~ as.factor(treatment) + as.factor(period), flair_data_lin_5)
summary(mod_lm_per)
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_lm_per <- lm(response ~ as.factor(treatment) + as.factor(period), flair_data_step_0.25)
summary(mod_lm_per)
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_lm_per <- lm(response ~ as.factor(treatment) + as.factor(period), flair_data_step_1.5)
summary(mod_lm_per)
```

### Stepwise trend, $\lambda=5$

```{r}
mod_lm_per <- lm(response ~ as.factor(treatment) + as.factor(period), flair_data_step_5)
summary(mod_lm_per)
```


## Regression model - calendar time adjustment

### Linear trend, $\lambda=0.25$

```{r}
mod_lm_cal <- lm(response ~ as.factor(treatment) + as.factor(cal_time), flair_data_lin_0.25)
summary(mod_lm_cal)
```

### Linear trend, $\lambda=1.5$

```{r}
mod_lm_cal <- lm(response ~ as.factor(treatment) + as.factor(cal_time), flair_data_lin_1.5)
summary(mod_lm_cal)
```

### Linear trend, $\lambda=5$

```{r}
mod_lm_cal <- lm(response ~ as.factor(treatment) + as.factor(cal_time), flair_data_lin_5)
summary(mod_lm_cal)
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_lm_cal <- lm(response ~ as.factor(treatment) + as.factor(cal_time), flair_data_step_0.25)
summary(mod_lm_cal)
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_lm_cal <- lm(response ~ as.factor(treatment) + as.factor(cal_time), flair_data_step_1.5)
summary(mod_lm_cal)
```

### Stepwise trend, $\lambda=5$

```{r}
mod_lm_cal <- lm(response ~ as.factor(treatment) + as.factor(cal_time), flair_data_step_5)
summary(mod_lm_cal)
```


## Mixed model - period adjustment

### Linear trend, $\lambda=0.25$

```{r}
mod_mix_per <- lmer(response ~ as.factor(treatment) + (1 | period), flair_data_lin_0.25)
summary(mod_mix_per)
```

### Linear trend, $\lambda=1.5$

```{r}
mod_mix_per <- lmer(response ~ as.factor(treatment) + (1 | period), flair_data_lin_1.5)
summary(mod_mix_per)
```

### Linear trend, $\lambda=5$

```{r}
mod_mix_per <- lmer(response ~ as.factor(treatment) + (1 | period), flair_data_lin_5)
summary(mod_mix_per)
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_mix_per <- lmer(response ~ as.factor(treatment) + (1 | period), flair_data_step_0.25)
summary(mod_mix_per)
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_mix_per <- lmer(response ~ as.factor(treatment) + (1 | period), flair_data_step_1.5)
summary(mod_mix_per)
```

### Stepwise trend, $\lambda=5$

```{r}
mod_mix_per <- lmer(response ~ as.factor(treatment) + (1 | period), flair_data_step_5)
summary(mod_mix_per)
```

## Mixed model - calendar time adjustment

### Linear trend, $\lambda=0.25$

```{r}
mod_mix_cal <- lmer(response ~ as.factor(treatment) + (1 | cal_time), flair_data_lin_0.25)
summary(mod_mix_cal)
```

### Linear trend, $\lambda=1.5$

```{r}
mod_mix_cal <- lmer(response ~ as.factor(treatment) + (1 | cal_time), flair_data_lin_1.5)
summary(mod_mix_cal)
```

### Linear trend, $\lambda=5$

```{r}
mod_mix_cal <- lmer(response ~ as.factor(treatment) + (1 | cal_time), flair_data_lin_5)
summary(mod_mix_cal)
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_mix_cal <- lmer(response ~ as.factor(treatment) + (1 | cal_time), flair_data_step_0.25)
summary(mod_mix_cal)
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_mix_cal <- lmer(response ~ as.factor(treatment) + (1 | cal_time), flair_data_step_1.5)
summary(mod_mix_cal)
```

### Stepwise trend, $\lambda=5$

```{r}
mod_mix_cal <- lmer(response ~ as.factor(treatment) + (1 | cal_time), flair_data_step_5)
summary(mod_mix_cal)
```

## Mixed model (AR1) - period adjustment

### Linear trend, $\lambda=0.25$

```{r}
mod_mix_ar1_per <- fitme(response ~ as.factor(treatment) + AR1(1 | period), flair_data_lin_0.25)

summary.HLfit(mod_mix_ar1_per)
res <- summary.HLfit(mod_mix_ar1_per, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_per, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Linear trend, $\lambda=1.5$

```{r}
mod_mix_ar1_per <- fitme(response ~ as.factor(treatment) + AR1(1 | period), flair_data_lin_1.5)

summary.HLfit(mod_mix_ar1_per)
res <- summary.HLfit(mod_mix_ar1_per, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_per, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Linear trend, $\lambda=5$

```{r}
mod_mix_ar1_per <- fitme(response ~ as.factor(treatment) + AR1(1 | period), flair_data_lin_5)

summary.HLfit(mod_mix_ar1_per)
res <- summary.HLfit(mod_mix_ar1_per, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_per, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_mix_ar1_per <- fitme(response ~ as.factor(treatment) + AR1(1 | period), flair_data_step_0.25)

summary.HLfit(mod_mix_ar1_per)
res <- summary.HLfit(mod_mix_ar1_per, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_per, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_mix_ar1_per <- fitme(response ~ as.factor(treatment) + AR1(1 | period), flair_data_step_1.5)

summary.HLfit(mod_mix_ar1_per)
res <- summary.HLfit(mod_mix_ar1_per, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_per, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Stepwise trend, $\lambda=5$

```{r}
mod_mix_ar1_per <- fitme(response ~ as.factor(treatment) + AR1(1 | period), flair_data_step_5)

summary.HLfit(mod_mix_ar1_per)
res <- summary.HLfit(mod_mix_ar1_per, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_per, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```


## Mixed model (AR1) - calendar time adjustment

### Linear trend, $\lambda=0.25$

```{r}
mod_mix_ar1_cal <- fitme(response ~ as.factor(treatment) + AR1(1 | cal_time), flair_data_lin_0.25)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Linear trend, $\lambda=1.5$

```{r}
mod_mix_ar1_cal <- fitme(response ~ as.factor(treatment) + AR1(1 | cal_time), flair_data_lin_1.5)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Linear trend, $\lambda=5$

```{r}
mod_mix_ar1_cal <- fitme(response ~ as.factor(treatment) + AR1(1 | cal_time), flair_data_lin_5)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_mix_ar1_cal <- fitme(response ~ as.factor(treatment) + AR1(1 | cal_time), flair_data_step_0.25)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_mix_ar1_cal <- fitme(response ~ as.factor(treatment) + AR1(1 | cal_time), flair_data_step_1.5)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```

### Stepwise trend, $\lambda=5$

```{r}
mod_mix_ar1_cal <- fitme(response ~ as.factor(treatment) + AR1(1 | cal_time), flair_data_step_5)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["as.factor(treatment)2", "t-value"]), eff_df)) # p-value
```



## Splines with knotes according to periods

### Linear trend, $\lambda=0.25$

```{r}
period_start <- c(max(which(flair_data_lin_0.25$period==1)), max(which(flair_data_lin_0.25$period==2)))

mod_splines_per <- lm(response ~ as.factor(treatment) + bs(j, knots = period_start, degree = 3), flair_data_lin_0.25)
summary(mod_splines_per)
```

### Linear trend, $\lambda=1.5$

```{r}
period_start <- c(max(which(flair_data_lin_1.5$period==1)), max(which(flair_data_lin_1.5$period==2)))

mod_splines_per <- lm(response ~ as.factor(treatment) + bs(j, knots = period_start, degree = 3), flair_data_lin_1.5)
summary(mod_splines_per)
```

### Linear trend, $\lambda=5$

```{r}
period_start <- c(max(which(flair_data_lin_5$period==1)), max(which(flair_data_lin_5$period==2)))

mod_splines_per <- lm(response ~ as.factor(treatment) + bs(j, knots = period_start, degree = 3), flair_data_lin_5)
summary(mod_splines_per)
```

### Stepwise trend, $\lambda=0.25$

```{r}
period_start <- c(max(which(flair_data_step_0.25$period==1)), max(which(flair_data_step_0.25$period==2)))

mod_splines_per <- lm(response ~ as.factor(treatment) + bs(j, knots = period_start, degree = 3), flair_data_step_0.25)
summary(mod_splines_per)
```

### Stepwise trend, $\lambda=1.5$

```{r}
period_start <- c(max(which(flair_data_step_1.5$period==1)), max(which(flair_data_step_1.5$period==2)))

mod_splines_per <- lm(response ~ as.factor(treatment) + bs(j, knots = period_start, degree = 3), flair_data_step_1.5)
summary(mod_splines_per)
```

### Stepwise trend, $\lambda=5$

```{r}
period_start <- c(max(which(flair_data_step_5$period==1)), max(which(flair_data_step_5$period==2)))

mod_splines_per <- lm(response ~ as.factor(treatment) + bs(j, knots = period_start, degree = 3), flair_data_step_5)
summary(mod_splines_per)
```


## Splines with knotes according to calendar time

### Linear trend, $\lambda=0.25$

```{r}
cal_time_start <- c()
for (i in unique(flair_data_lin_0.25$cal_time)) {
  cal_time_start <- c(cal_time_start, max(flair_data_lin_0.25[flair_data_lin_0.25$cal_time==i,]$j))
}

mod_splines_cal <- lm(response ~ as.factor(treatment) + bs(j, knots = cal_time_start[1:length(cal_time_start)-1], degree = 3), flair_data_lin_0.25)
summary(mod_splines_cal)
```

### Linear trend, $\lambda=1.5$

```{r}
cal_time_start <- c()
for (i in unique(flair_data_lin_1.5$cal_time)) {
  cal_time_start <- c(cal_time_start, max(flair_data_lin_1.5[flair_data_lin_1.5$cal_time==i,]$j))
}

mod_splines_cal <- lm(response ~ as.factor(treatment) + bs(j, knots = cal_time_start[1:length(cal_time_start)-1], degree = 3), flair_data_lin_1.5)
summary(mod_splines_cal)
```

### Linear trend, $\lambda=5$

```{r}
cal_time_start <- c()
for (i in unique(flair_data_lin_5$cal_time)) {
  cal_time_start <- c(cal_time_start, max(flair_data_lin_5[flair_data_lin_5$cal_time==i,]$j))
}

mod_splines_cal <- lm(response ~ as.factor(treatment) + bs(j, knots = cal_time_start[1:length(cal_time_start)-1], degree = 3), flair_data_lin_5)
summary(mod_splines_cal)
```

### Stepwise trend, $\lambda=0.25$

```{r}
cal_time_start <- c()
for (i in unique(flair_data_step_0.25$cal_time)) {
  cal_time_start <- c(cal_time_start, max(flair_data_step_0.25[flair_data_step_0.25$cal_time==i,]$j))
}

mod_splines_cal <- lm(response ~ as.factor(treatment) + bs(j, knots = cal_time_start[1:length(cal_time_start)-1], degree = 3), flair_data_step_0.25)
summary(mod_splines_cal)
```

### Stepwise trend, $\lambda=1.5$

```{r}
cal_time_start <- c()
for (i in unique(flair_data_step_1.5$cal_time)) {
  cal_time_start <- c(cal_time_start, max(flair_data_step_1.5[flair_data_step_1.5$cal_time==i,]$j))
}

mod_splines_cal <- lm(response ~ as.factor(treatment) + bs(j, knots = cal_time_start[1:length(cal_time_start)-1], degree = 3), flair_data_step_1.5)
summary(mod_splines_cal)
```

### Stepwise trend, $\lambda=5$

```{r}
cal_time_start <- c()
for (i in unique(flair_data_step_5$cal_time)) {
  cal_time_start <- c(cal_time_start, max(flair_data_step_5[flair_data_step_5$cal_time==i,]$j))
}

mod_splines_cal <- lm(response ~ as.factor(treatment) + bs(j, knots = cal_time_start[1:length(cal_time_start)-1], degree = 3), flair_data_step_5)
summary(mod_splines_cal)
```


## Pooled analysis

### Linear trend, $\lambda=0.25$

```{r}
mod_pool <- lm(response ~ as.factor(treatment), flair_data_lin_0.25 %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)
```

### Linear trend, $\lambda=1.5$

```{r}
mod_pool <- lm(response ~ as.factor(treatment), flair_data_lin_1.5 %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)
```

### Linear trend, $\lambda=5$

```{r}
mod_pool <- lm(response ~ as.factor(treatment), flair_data_lin_5 %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_pool <- lm(response ~ as.factor(treatment), flair_data_step_0.25 %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_pool <- lm(response ~ as.factor(treatment), flair_data_step_1.5 %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)
```

### Stepwise trend, $\lambda=5$

```{r}
mod_pool <- lm(response ~ as.factor(treatment), flair_data_step_5 %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)
```


## Separate analysis

### Linear trend, $\lambda=0.25$

```{r}
mod_sep <- lm(response ~ as.factor(treatment), flair_data_lin_0.25 %>% filter(treatment %in% c(0, 2), period %in% c(2, 3)))
summary(mod_sep)
```

### Linear trend, $\lambda=1.5$

```{r}
mod_sep <- lm(response ~ as.factor(treatment), flair_data_lin_1.5 %>% filter(treatment %in% c(0, 2), period %in% c(2, 3)))
summary(mod_sep)
```

### Linear trend, $\lambda=5$

```{r}
mod_sep <- lm(response ~ as.factor(treatment), flair_data_lin_5 %>% filter(treatment %in% c(0, 2), period %in% c(2, 3)))
summary(mod_sep)
```

### Stepwise trend, $\lambda=0.25$

```{r}
mod_sep <- lm(response ~ as.factor(treatment), flair_data_step_0.25 %>% filter(treatment %in% c(0, 2), period %in% c(2, 3)))
summary(mod_sep)
```

### Stepwise trend, $\lambda=1.5$

```{r}
mod_sep <- lm(response ~ as.factor(treatment), flair_data_step_1.5 %>% filter(treatment %in% c(0, 2), period %in% c(2, 3)))
summary(mod_sep)
```

### Stepwise trend, $\lambda=5$

```{r}
mod_sep <- lm(response ~ as.factor(treatment), flair_data_step_5 %>% filter(treatment %in% c(0, 2), period %in% c(2, 3)))
summary(mod_sep)
```







# Bias simulations

```{r}
n_sim <- 10000    
set.seed(123)

res_fix_step_0.25 <- c()
for (i in 1:n_sim) {
  res_fix_step_0.25 <- c(res_fix_step_0.25, fixmodel_cont(sim_flair(lambda = 0.25, trend_pattern = "stepwise"), 2)$treat_effect - (-log(0.13)))
  if (i %in% seq(1000, 10000, by=1000)) { # keep track of the progress
    print(i) 
  }
}

t.test(res_fix_step_0.25)



res_splines_step_0.25 <- c()
for (i in 1:n_sim) {
  res_splines_step_0.25 <- c(res_splines_step_0.25, splines_cont(sim_flair(lambda = 0.25, trend_pattern = "stepwise"), 2)$treat_effect - (-log(0.13)))
  if (i %in% seq(1000, 10000, by=1000)) { # keep track of the progress
    print(i) 
  }
}

t.test(res_splines_step_0.25)



res_mix_step_0.25 <- c()
for (i in 1:n_sim) {
  res_mix_step_0.25 <- c(res_mix_step_0.25, mixmodel_cont(sim_flair(lambda = 0.25, trend_pattern = "stepwise"), 2)$treat_effect - (-log(0.13)))
  if (i %in% seq(1000, 10000, by=1000)) { # keep track of the progress
    print(i) 
  }
}

t.test(res_mix_step_0.25)



res_mix_ar1_step_0.25 <- c()
for (i in 1:n_sim) {
  res_mix_ar1_step_0.25 <- c(res_mix_ar1_step_0.25, mixmodel_AR1_cont(sim_flair(lambda = 0.25, trend_pattern = "stepwise"), 2)$treat_effect - (-log(0.13)))
  if (i %in% seq(1000, 10000, by=1000)) { # keep track of the progress
    print(i) 
  }
}

t.test(res_mix_ar1_step_0.25)
```


## Figure S11

```{r, fig.height=6, fig.width=7}
bias_res <- data.frame(model = c(rep("Fixed effect model", n_sim),
                                 rep("Spline regression", n_sim),
                                 rep("Mixed model", n_sim),
                                 rep("Mixed model (AR1)", n_sim)),
                       bias = c(res_fix_step_0.25,
                                res_splines_step_0.25,
                                res_mix_step_0.25,
                                res_mix_ar1_step_0.25))
  
  
ggplot(bias_res) +
  geom_histogram(aes(bias), color="darkred", fill="darkred", alpha=0.6, bins = 40) +
  facet_wrap(~ model, nrow=2) +
  labs(x = "Estimation error", y = "Count") +
  geom_vline(xintercept = 0, color = "black") +
  theme_bw(base_size = 13.8)

ggsave("figures/FLAIR_bias.png", width = 7, height = 6)
ggsave("figures/FLAIR_bias.pdf", width = 7, height = 6)
ggsave("figures/FLAIR_bias.tiff", width = 7, height = 6)
```







