---
title: "Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls"
subtitle: "Case study: PSP data (synthetic)"
author: "Pavla Krotka, Martin Posch, Mohamed Gewily, Günter Höglinger, Marta Bofill Roig"
date: "2024"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show 
    theme: flatly
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(collapse = TRUE,
               echo = TRUE,
               comment = "#>",
               warning = FALSE,
               message = FALSE,
               fig.align = "center",
               out.width = "100%")
```

```{r, include=FALSE}
library(kableExtra)
library(latex2exp)
library(tidyverse)
library(splines)
library(lmerTest)
library(spaMM)
library(viridis)

set.seed(717)
```


# Introduction

This file contains all code to generate synthetic data in order to mimic the case study using PSP data presented in Section 5 of the paper *"Statistical modeling to adjust for time trends in adaptive platform trials utilizing non-concurrent controls"* by Pavla Krotka, Martin Posch, Mohamed Gewily, Günter Höglinger, Marta Bofill Roig.


# Generation of synthetic data

```{r}
# Generate random values from the Beta distribution for patient recruitment
set.seed(22)
beta_values <- rbeta(315, 3, 1.5)

# Scale the values to the desired range (1 to 742)
recruitment <- round(1 + beta_values * (742 - 1))

# Generate randomized allocations and patient responses
set.seed(13)
period_1 <- sample(c(rep(0, 55), rep(1, 61)))
period_2 <- sample(c(rep(0, 68), rep(1, 65), rep(2, 66)))

resp_1 <- rnorm(n = 116, mean = 38.6, sd = 11)
resp_2 <- rnorm(n = 199, mean = 34.9, sd = 11)

# Create dataframe with trial data
data_case_study <- data.frame(recruitment = sort(recruitment),
                              response_visit_1 = c(resp_1, resp_2),
                              treatment = as.factor(c(period_1, period_2)),
                              period = c(rep(1, 116), rep(2, 199)),
                              cal_time = factor(ceiling(recruitment/90), levels = c(1:9)))
```



# Abbvie trial data - baseline measurement

## Sample sizes per treatment arm

```{r}
table(data_case_study$treatment)
```

## Sample size per period

```{r}
table(data_case_study$period)
```


## Sample size per treatment arm and period

```{r}
table(data_case_study$treatment, data_case_study$period)
```

## Sample sizes per calendar time interval

```{r}
table(data_case_study$cal_time)
```


## Sample sizes per treatment arm and calendar time interval

```{r}
table(data_case_study$treatment, data_case_study$cal_time)
```


## Time trend

```{r}
test_trend <- lm(response_visit_1 ~ treatment + recruitment, data_case_study)
summary(test_trend)
```

# Figure 13


```{r}
per2_start <- max(data_case_study[data_case_study$period==1,]$recruitment)

ggplot(data_case_study) +
  geom_point(aes(recruitment, response_visit_1, color = treatment)) +
  geom_line(aes(recruitment, fitted(test_trend), color = treatment), size = 1) +
  labs(x = "Trial duration in days", y = "PSPRS score at baseline", color = "Treatment:") +
  scale_color_viridis(discrete = T) +
  geom_vline(xintercept = 90*c(0:9), linetype = "dashed", color = "gray40") +
  geom_vline(xintercept = per2_start, linetype = "dashed", color = "darkred", size = 1) +
  theme_bw() +
  theme(legend.position = "bottom")
```

# Analysis approaches

## Regression model - period adjustment

```{r}
# Regression model - period adjustment
mod_lm_per <- lm(response_visit_1 ~ treatment + period, data_case_study)
summary(mod_lm_per)

mod_lm_per_est <- summary(mod_lm_per)$coefficients["treatment2", "Estimate"]
mod_lm_per_se <- summary(mod_lm_per)$coefficients["treatment2", "Std. Error"]
mod_lm_per_pval <- summary(mod_lm_per)$coefficients["treatment2", "Pr(>|t|)"]
```

## Regression model - calendar time adjustment

```{r}
# Regression model - calendar time adjustment
mod_lm_cal <- lm(response_visit_1 ~ treatment + cal_time, data_case_study)
summary(mod_lm_cal)

mod_lm_cal_est <- summary(mod_lm_cal)$coefficients["treatment2", "Estimate"]
mod_lm_cal_se <- summary(mod_lm_cal)$coefficients["treatment2", "Std. Error"]
mod_lm_cal_pval <- summary(mod_lm_cal)$coefficients["treatment2", "Pr(>|t|)"]
```

## Mixed model - calendar time adjustment

```{r}
# Mixed model - calendar time adjustment
mod_mix_cal <- lmer(response_visit_1 ~ treatment + (1 | cal_time), data_case_study)
summary(mod_mix_cal)

mod_mix_cal_est <- summary(mod_mix_cal)$coefficients["treatment2", "Estimate"]
mod_mix_cal_se <- summary(mod_mix_cal)$coefficients["treatment2", "Std. Error"]
mod_mix_cal_pval <- summary(mod_mix_cal)$coefficients["treatment2", "Pr(>|t|)"]
```


## Mixed model (AR1) - calendar time adjustment

```{r}
# Mixed model - calendar time adjustment
mod_mix_ar1_cal <- fitme(response_visit_1 ~ treatment + AR1(1 | cal_time), data_case_study)

summary.HLfit(mod_mix_ar1_cal)
res <- summary.HLfit(mod_mix_ar1_cal, verbose = FALSE)

IC <- get_any_IC(mod_mix_ar1_cal, verbose = FALSE)
IC
eff_df <- IC["       effective df:"] # effective degrees of freedom

2*(1-pt(abs(res$beta_table["treatment2", "t-value"]), eff_df)) # p-value

mod_mix_ar1_cal_est <- summary(mod_mix_ar1_cal, verbose = F)$beta_table["treatment2", "Estimate"]
mod_mix_ar1_cal_se <- summary(mod_mix_ar1_cal, verbose = F)$beta_table["treatment2", "Cond. SE"]
mod_mix_ar1_cal_pval <- 2*(1-pt(abs(res$beta_table["treatment2", "t-value"]), eff_df))
```

## Spline regression - period adjustment

```{r}
# Spline regression - period adjustment
per2_start <- max(data_case_study[data_case_study$period==1,]$recruitment)

mod_splines_per <- lm(response_visit_1 ~ treatment + bs(recruitment, knots = per2_start, degree = 3), data_case_study)
summary(mod_splines_per)

mod_splines_per_est <- summary(mod_splines_per)$coefficients["treatment2", "Estimate"]
mod_splines_per_se <- summary(mod_splines_per)$coefficients["treatment2", "Std. Error"]
mod_splines_per_pval <- summary(mod_splines_per)$coefficients["treatment2", "Pr(>|t|)"]
```

## Spline regression - calendar time adjustment

```{r}
# Spline regression - calendar time adjustment
cal_time_start <- c()
for (i in unique(data_case_study$cal_time)) {
  cal_time_start <- c(cal_time_start, max(data_case_study[data_case_study$cal_time==i,]$recruitment))
}

cal_time_start <- cal_time_start[-length(cal_time_start)]

mod_splines_cal <- lm(response_visit_1 ~ treatment + bs(recruitment, knots = cal_time_start, degree = 3), data_case_study)
summary(mod_splines_cal)

mod_splines_cal_est <- summary(mod_splines_cal)$coefficients["treatment2", "Estimate"]
mod_splines_cal_se <- summary(mod_splines_cal)$coefficients["treatment2", "Std. Error"]
mod_splines_cal_pval <- summary(mod_splines_cal)$coefficients["treatment2", "Pr(>|t|)"]
```

## Pooled analysis

```{r}
# Pooled analysis
mod_pool <- lm(response_visit_1 ~ treatment, data_case_study %>% filter(treatment %in% c(0, 2)))
summary(mod_pool)

mod_pool_est <- summary(mod_pool)$coefficients["treatment2", "Estimate"]
mod_pool_se <- summary(mod_pool)$coefficients["treatment2", "Std. Error"]
mod_pool_pval <- summary(mod_pool)$coefficients["treatment2", "Pr(>|t|)"]
```

## Separate analysis

```{r}
# Separate analysis
mod_sep <- lm(response_visit_1 ~ treatment, data_case_study %>% filter(treatment %in% c(0, 2), period==2))
summary(mod_sep)

mod_sep_est <- summary(mod_sep)$coefficients["treatment2", "Estimate"]
mod_sep_se <- summary(mod_sep)$coefficients["treatment2", "Std. Error"]
mod_sep_pval <- summary(mod_sep)$coefficients["treatment2", "Pr(>|t|)"]
```


# TABLE 2


```{r}
table_2 <- data.frame(`Analysis approach` = c("Fixed effect model", "Fixed effect model", "Mixed model", 
                                              "Mixed model (AR1)", "Spline regression", "Spline regression", "Pooled analysis", "Separate analysis"),
                      Adjustment = c("Periods", rep("Calendar time units", 3), "Periods", "Calendar time units", rep("-", 2)),
                      `Effect estimate` = round(c(mod_lm_per_est, mod_lm_cal_est, mod_mix_cal_est, mod_mix_ar1_cal_est, mod_splines_per_est, mod_splines_cal_est, mod_pool_est, mod_sep_est), 3),
                      `Std. error` = round(c(mod_lm_per_se, mod_lm_cal_se, mod_mix_cal_se, mod_mix_ar1_cal_se, mod_splines_per_se, mod_splines_cal_se, mod_pool_se, mod_sep_se), 3),
                      `p-value` = round(c(mod_lm_per_pval, mod_lm_cal_pval, mod_mix_cal_pval, mod_mix_ar1_cal_pval, mod_splines_per_pval, mod_splines_cal_pval, mod_pool_pval, mod_sep_pval), 3),
                      
                      check.names = F)

kable(table_2, booktabs = T) %>%
  kable_styling(bootstrap_options = "striped")
```







